### 2015-03-08 paper_10_intro.tex
added abbreviations page

    This work 
    provides a technique for a Software Development Organization (SDO) to create
    a single number score which indicates the overall performance
    of the organization. The
    score is based upon data collected for 5 key result indicators
    of an SDO: quality, availability,
    satisfaction, schedule, and requirements.  It is not meant
    to be comparative between organizations, but to 
    form a historical baseline for a specific organization.  The single
    number is targeted for upper-level management whom need a quick
    and simple strategy to evaluate the performance of the organization.

@@

    SDOs need a framework to define
    what measurements should be tracked and how those measurements
    should be reported.  The Cumulative Result Indicator (CRI)
    framework provides a solid foundation
    for a consistent evaluation.  CRI analyzes the historical performance
    of an SDO
    to create a baseline in order to provide a broad
    view of the overall organization.  It is common for software development
    organizations to measure and focus soley on the source code
    being produced.  However, an SDO does more
    than just produce source code.  There is documentation to be
    written, testcases to be created, systems to be deployed, and
    decisions to be made.  The framework provides an evaluation
    of the overall SDO, not just the
    source code.

@@

        \item[Application] - a software system or a collection of other applications 
        \item[Release] - A collection of projects being put in production on a specified date
        \item[Project] - A body of work involving zero or more applications in preparation for a release
        \item[SIT]   (Systems Integration Testing) - The initial step of testing
            after the development phase of the SDLC.  This is typically 
            performed by members of the SDO.  It is validation that all the
            software components function together as expected.
        \item[UAT]   (User Acceptance Testing) - The final step of testing
            when a select few members of the user group are invited to
            validate the software system. Once validation has occured for
            UAT, the software system is ready to proceed to production
        \item[PROD]  (Production) - The software has been released to the final audience.
        \item[defect] ``A software defect is a bug or error that causes
                    software to either stop operating or to produce
                    invalid or unacceptable results'' as quoted from Capers Jones
                \cite{Jones2009}.  It is important to mention that even though defects are typically 
                found in the computer code, a defect should not be isolated to just code.   
                A poorly written requirement or missed test cases can 
                both be considered a defect.  Other common names for a defect 
                are: bug, error, fault, or ticket.
    \end{description}

\subsection{CMMI}
    The Capability Maturity Model Integration (CMMI) is one of the most 
    widely acknowledged models for process 
    improvement in software development.  CMMI offers a generic guideline 
    and appraisal program for process 
    improvement.  It was created and is administered by the 
    Software Engineering Institute at
    Carnegie Mellon University \cite{CMMI}. While the CMMI is not specific to software development,
    it is often applied in software development settings.
    CMMI certification is required for many United States Government 
    and Department of Defense contracts.

@@

    Software Engineering Method and Theory (SEMAT) is claiming to 
    be the ``new software engineering'' 
    \cite{Jacobson2014}.  The authors rightfully claim that 
    software engineering lacks
    an underlying theoretical foundation found in other 
    engineering disciplines.  This lack of theory
    has led software engineering to not be engineering, but 
    rather a craft.  The goal of 
    SEMAT is to merge the craftsmanship and engineering to 
    provide a foundation for software
    engineering.  The primary initiative of SEMAT 
    has been the creation of a kernel for 
    software engineering.  The kernel is the minimal set 
    of things common to all software development
    endeavours. The three parts to the kernel are:
    \begin{description}
        \item[Measurement] - There must exist a means to 
            determine the health and progress of an endeavour
        \item[Categorization]-  The activities must progress 
            through categories during an endeavour
        \item[Competencies] - Specific competencies will 
            be required for completing activities
    \end{description}

>>>>

    This work 
    provides a technique for a Software Development Organization (SDO)
    \nomenclature{SDO}{Software Development Organization}to create
    a single number score which indicates the overall performance
    of the organization. The
    score is based upon data collected for 5 key result indicators
    of an SDO: quality, availability,
    satisfaction, schedule, and requirements.  It is not meant
    to be comparative between organizations, but to 
    form a historical baseline for a specific organization.  The single
    number is targeted for upper-level management whom need a quick
    and simple strategy to evaluate the performance of the organization.

@@

    SDOs need a framework to define
    what measurements should be tracked and how those measurements
    should be reported.  The Cumulative Result Indicator (CRI)
    \nomenclature{CRI}{Cumulative Result Indicator}
    framework provides a solid foundation
    for a consistent evaluation.  CRI analyzes the historical performance
    of an SDO
    to create a baseline in order to provide a broad
    view of the overall organization.  It is common for software development
    organizations to measure and focus soley on the source code
    being produced.  However, an SDO does more
    than just produce source code.  There is documentation to be
    written, testcases to be created, systems to be deployed, and
    decisions to be made.  The framework provides an evaluation
    of the overall SDO, not just the
    source code.

@@

        \item[Application] - a software system or a collection of other applications 
        \item[Release] - A collection of projects being put in production on a specified date
        \item[Project] - A body of work involving zero or more applications in preparation for a release 
        \item[SIT] (Systems Integration Testing) - The initial step of testing
            after the development phase of the SDLC.  This is typically 
            performed by members of the SDO.  It is validation that all the
            software components function together as expected.
            \nomenclature{SIT}{Systems Integration Testing} 
        \item[UAT] (User Acceptance Testing) - The final step of testing
            when a select few members of the user group are invited to
            validate the software system. Once validation has occured for
            UAT, the software system is ready to proceed to production
            \nomenclature{UAT}{User Acceptance Testing} 
        \item[PROD] (Production) - The software has been released to the 
        final audience. \nomenclature{PROD}{Production} 
        \item[defect] ``A software defect is a bug or error that causes
                    software to either stop operating or to produce
                    invalid or unacceptable results'' as quoted from Capers Jones
                \cite{Jones2009}.  It is important to mention that even though defects are typically 
                found in the computer code, a defect should not be isolated to just code.   
                A poorly written requirement or missed test cases can 
                both be considered a defect.  Other common names for a defect 
                are: bug, error, fault, or ticket.
    \end{description}

\subsection{CMMI}
    The Capability Maturity Model Integration (CMMI)
    \nomenclature{CMMI}{Capability Maturity Model Integration} is one of the most 
    widely acknowledged models for process 
    improvement in software development.  CMMI offers a generic guideline 
    and appraisal program for process 
    improvement.  It was created and is administered by the 
    Software Engineering Institute at
    Carnegie Mellon University \cite{CMMI}. While the CMMI is not specific to software development,
    it is often applied in software development settings.
    CMMI certification is required for many United States Government 
    and Department of Defense contracts.

@@

    Software Engineering Method and Theory (SEMAT) is claiming to 
    be the ``new software engineering'' 
    \cite{Jacobson2014}. 
    \nomenclature{SEMAT}{Software Engineering Method and Theory}
    The authors rightfully claim that 
    software engineering lacks
    an underlying theoretical foundation found in other 
    engineering disciplines.  This lack of theory
    has led software engineering to not be engineering, but 
    rather a craft.  The goal of 
    SEMAT is to merge the craftsmanship and engineering to 
    provide a foundation for software
    engineering.  The primary initiative of SEMAT 
    has been the creation of a kernel for 
    software engineering.  The kernel is the minimal set 
    of things common to all software development
    endeavours. The three parts to the kernel are:
    \begin{description}
        \item[Measurement] - There must exist a means to 
            determine the health and progress of an endeavour
        \item[Categorization]-  The activities must progress 
            through categories during an endeavour
        \item[Competencies] - Specific competencies will 
            be required for completing activities
    \end{description}

### 2015-03-08 paper_20_sdo.tex
added abbreviations page

\section{A SOFTWARE DEVELOPMENT ORGANIZATION (SDO)}
A \textit{Software Development Organization (SDO)} is any organization or subset of an organization that is responsible
for the creation, deployment, and maintenance of software.  Many times an SDO
is a company that produces software.  Other times, an SDO is contained
within the Information Technology department of a larger organization. Some of the job roles
with an SDO are: software engineer, system administrator, software quality analyst,
programmer, database administrator, and documentation specialist.

@@

\subsection{THE SOFTWARE DEVELOPMENT LIFE CYCLE}
    The discipline of software engineering has created a workflow for developing
    software.  
    This workflow is called the \textit{Software Development Life Cycle (SDLC)}.
    SDLC can be defined as \cite{Ruparelia2010}:
    \begin{quote}
     [...] a conceptual framework or process that considers the structure of the stages
     involved in the development of an application from its initial feasibility study
     through to its deployment in the field and maintenance.
    \end{quote}
    While the SDLC states what needs to be done, there are numerous models 
    that formalize exactly how to perform the SDLC.  The models contain
    steps that are commonly referred to as a phases. A few of the popular
    models are described below.

@@

\subsection{WHAT IS SOFTWARE ENGINEERING?}
\label{sec:swe}
    \textit{Software Engineering} as a term dates back to the 
    1968 North Atlantic Treaty Organization (NATO) conference \cite{Tsui2013, Naur1969}. Over the 
    years many definitions have been provided.  IEEE provides a definition that encompasses many of the
    other definitions.  IEEE ISO610.12 defines software engineering as, 
    ''The application of a systematic, disciplined, quantifiable approach to the development, operation, and maintenance of software''
    \cite{Ieee1990}.

>>>>

\section{A SOFTWARE DEVELOPMENT ORGANIZATION}
A \textit{Software Development Organization (SDO)} is any organization or subset of an organization that is responsible
for the creation, deployment, and maintenance of software.  Many times an SDO
is a company that produces software.  Other times, an SDO is contained
within the Information Technology department of a larger organization. Some of the job roles
with an SDO are: software engineer, system administrator, software quality analyst,
programmer, database administrator, and documentation specialist.

@@

\subsection{THE SOFTWARE DEVELOPMENT LIFE CYCLE}
    The discipline of software engineering has created a workflow for developing
    software.  
    This workflow is called the \textit{Software Development Life Cycle (SDLC)}
    \nomenclature{SDLC}{Software Development Life Cycle}.
    SDLC can be defined as \cite{Ruparelia2010}:
    \begin{quote}
     [...] a conceptual framework or process that considers the structure of the stages
     involved in the development of an application from its initial feasibility study
     through to its deployment in the field and maintenance.
    \end{quote}
    While the SDLC states what needs to be done, there are numerous models 
    that formalize exactly how to perform the SDLC.  The models contain
    steps that are commonly referred to as a phases. A few of the popular
    models are described below.

@@

\subsection{WHAT IS SOFTWARE ENGINEERING?}
\label{sec:swe}
    \textit{Software Engineering} as a term dates back to the 
    1968 North Atlantic Treaty Organization (NATO) conference \cite{Tsui2013, Naur1969}. Over the 
    years many definitions have been provided.  Institute of Electrical 
    and Electronics Engineers (IEEE) provides a definition that encompasses 
    many of the other definitions.
    \nomenclature{IEEE}{Institute of Electrical and Electronics Engineers}
    IEEE ISO610.12 defines software engineering as, 
    ''The application of a systematic, disciplined, quantifiable approach to the development, operation, and maintenance of software''
    \cite{Ieee1990}.

### 2015-03-08 paper_30_measure.tex
added abbreviations page

    \subsection{METRICS}
        A \textit{metric} can be defined as a means of telling a complete story
        for the purpose of improving something \cite{Klubeck2011}.  Metrics are
        frequently indirect measurements and
        are very common in the measurement of SDOs.  
        The following are some examples of metrics that can be collected 
        for an SDO. 
        \begin{itemize}
            \item SLOC - The number of Source Lines of Code 
            \item NOM - The Number of Methods per class
            \item Complexity - A numerical measure of the code complexity,
                some common examples are McCabe \cite{McCabe1976} and 
                Halstead \cite{Halstead1977}
            \item Design - The amount of coupling and cohesion present 
                in the software code
            \item Source Code Analysis - Tools that determine whether 
                code adheres to specified set of rules. Common 
                examples are PMD\footnote{PMD is a source code analysis product.  
                It is not an acronym.} and FindBugs$^{TM}$ \cite{PMD, Findbugs}.
        \end{itemize}
        All these metrics are beneficial, but none of them tell the story of the entire
        SDO.  Most of the metrics, as seen in the list above, for an SDO focus on 
        the source code and development phase. 
        Since metrics are indirect, it can be very difficult match SDO performance with
        a metric or series of metrics.  Metrics are great for tracking, but decision making
        based upon metrics alone is difficult.  That is why many of the other techniques
        build upon metrics to provide a more complete overall picture of an SDO. Metrics are great starting point, but more is needed to properly 
        evaluate performance.

    \subsection{INDICATORS}
    \label{sec:indicators}
        Another common measurement technique is indicators. An
        \textit{indicator} is simply a performance measure. 
        Typically, a number of indicators will be placed
        together and displayed in some report or on some
        dashboard. 
        Indicators can be crucial measurements
        within any business setting, and an SDO is no exception. Determining
        the correct indicators for an organization can
        be difficult, and many organizations incorrectly classify
        the indicators \cite{parmenter2010}.  
        The differences between the indicators will be explored and
        possible measures for each indicator in an SDO will be
        presented.   The four
        types of indicators important to an SDO are shown
        in Table \ref{tab:indicators}.

@@

        \subsubsection{KEY PERFORMANCE INDICATOR (KPI) FOR AN SDO}
            \textit{Key Performance Indicators} are performance measures
            focusing on critical aspects for current and future organizational
            success.  Notice, KPIs are not focused on historical performance,
            and they clearly indicate how to drastically increase performance.
            KPIs allow a team to monitor current performance and 
            quickly take action to correct future performance.
            KPIs cover a shorter time frame than KRIs.
            KPIs consist of the following 7 characteristics \cite{parmenter2010}.
            \begin{enumerate}
                \item Not financial
                \item Measured frequently (hourly, daily, weekly)
                \item Acted on by CEO\footnote{Chief Executive Officer} 
                    and/or senior management
                \item Clearly indicate the action required
                \item Tie responsibility to a particular team
                \item Have a significant impact
                \item Encourage appropriate action
            \end{enumerate}

>>>>

    \subsection{METRICS}
        A \textit{metric} can be defined as a means of telling a complete story
        for the purpose of improving something \cite{Klubeck2011}.  Metrics are
        frequently indirect measurements and
        are very common in the measurement of SDOs.  
        The following are some examples of metrics that can be collected 
        for an SDO. 
        \begin{itemize}
            \item SLOC - The number of Source Lines of Code 
                \nomenclature{SLOC}{Source Lines Of Code}
            \item NOM - The Number of Methods per class
                \nomenclature{NOM}{Number of Methods}
            \item Complexity - A numerical measure of the code complexity,
                some common examples are McCabe \cite{McCabe1976} and 
                Halstead \cite{Halstead1977}
            \item Design - The amount of coupling and cohesion present 
                in the software code
            \item Source Code Analysis - Tools that determine whether 
                code adheres to specified set of rules. Common 
                examples are PMD\footnote{PMD is a source code analysis product.  
                It is not an acronym.} and FindBugs$^{TM}$ \cite{PMD, Findbugs}.
        \end{itemize}
        All these metrics are beneficial, but none of them tell the story of the entire
        SDO.  Most of the metrics, as seen in the list above, for an SDO focus on 
        the source code and development phase. 
        Since metrics are indirect, it can be very difficult match SDO performance with
        a metric or series of metrics.  Metrics are great for tracking, but decision making
        based upon metrics alone is difficult.  That is why many of the other techniques
        build upon metrics to provide a more complete overall picture of an SDO. Metrics are great starting point, but more is needed to properly 
        evaluate performance.

    \subsection{INDICATORS}
    \label{sec:indicators}
        Another common measurement technique is indicators. An
        \textit{indicator} is simply a performance measure. 
        Typically, a number of indicators will be placed
        together and displayed in some report or on some
        dashboard. 
        Indicators can be crucial measurements
        within any business setting, and an SDO is no exception. Determining
        the correct indicators for an organization can
        be difficult, and many organizations incorrectly classify
        the indicators \cite{parmenter2010}.  
        The differences between the indicators will be explored and
        possible measures for each indicator in an SDO will be
        presented.   The four
        categories of indicators important to an SDO are shown
        in Table \ref{tab:indicators}.

        \nomenclature{RI}{Result Indicator}
        \nomenclature{KRI}{Key Result Indicator}
        \nomenclature{PI}{Performance Indicator}
        \nomenclature{KPI}{Key Performance Indicator}

@@

        \subsubsection{KEY PERFORMANCE INDICATOR (KPI) FOR AN SDO}
            \textit{Key Performance Indicators} are performance measures
            focusing on critical aspects for current and future organizational
            success.  Notice, KPIs are not focused on historical performance,
            and they clearly indicate how to drastically increase performance.
            KPIs allow a team to monitor current performance and 
            quickly take action to correct future performance.
            KPIs cover a shorter time frame than KRIs.
            KPIs consist of the following 7 characteristics \cite{parmenter2010}.
            \begin{enumerate}
                \item Not financial
                \item Measured frequently (hourly, daily, weekly)
                \item Acted on by CEO\footnote{Chief Executive Officer} 
                    and/or upper-level management 
                    \nomenclature{CEO}{Chief Executive Officer}
                \item Clearly indicate the action required
                \item Tie responsibility to a particular team
                \item Have a significant impact
                \item Encourage appropriate action
            \end{enumerate}

### 2015-03-08 paper_40_cri.tex
added abbreviations page

    SDOs struggle to measure overall performance.  
    The \textit{Cumulative Result Indicator (CRI)} is an algorithm to provide a single 
    number score to measure the performance of an SDO. 
    It works by statistically analyzing the past performance of the 
    organization and using that information to score an organization 
    on current performance.  CRI is a collection of the following 5 
    elements, which are actually KRIs,
    for an SDO.
    \textbf{
    \begin{enumerate}
        \item Quality
        \item Availability
        \item Satisfaction
        \item Schedule
        \item Requirements
    \end{enumerate}
    }
    A separate CRI
    score is calculated for each element and 
    then aggregated together to form an overall 
    CRI score.  A new CRI score will be calculated based
    upon the selection of a given time period (weekly, monthly, quarterly).
    \textbf{CRI is not meant to be comparative between
    organizations, but to measure the amount of increase or decrease
    a single organization exhibits across elements}.  CRI is made to be easily 
    expandable to other elements if desired.

@@

                The development and testing effort can come from any of 
                the following choices for effort. It is possible
                that other measures will work for effort.
                \begin{description}
                    \item[Actual Time] This number is a representation of the 
                        total amount of time spent on a project.  This number
                        can be measured in any unit of time: hours, days, weeks,
                        etc.  Actual time can be applied to development or testing
                        effort.
                    \item[Estimated Time] This number is a representation of the 
                        initial estimated amount of time spent on a project.  This number
                        can be measured in any unit of time: hours, days, weeks,
                        etc.  Estimated time can be applied to development or testing
                        effort.  It is common for the estimated and actual times
                        to be diffent.
                    \item[Source Lines Of Code (SLOC)]  This number is the count of 
                        the total number of lines of source code for a project. 
                        Obviously, 
                        this item only counts as a level of
                        effort for development unless coding in used to generate
                        automated test cases. \footnote{Automated testing is the
                        process of creating software to automatically run tests
                        against other software.  The adoption of automated testing
                        is varied and it is not a solution in all cases 
                        \cite{Ramler2006}. }
                    \item[Modified Lines Of Code] This number is a count of the number 
                        of modified lines of source code. Modified lines is 
                        defined as the number of deleted, added, and modified 
                        lines of source code.  This number is different from above
                        since it does not include all the lines of source code.  Similar
                        to above, this number makes more sense for development effort.
                    \item[Test Cases] A test case is a step or series of steps followed
                        to validate some expected outcome of software.  Organizations
                        will create a number of test cases to be validated for a
                        software system.  The number of such test cases could
                        be used as a level of testing effort.
                \end{description}

>>>>

    SDOs struggle to measure overall performance.  
    The \textit{Cumulative Result Indicator (CRI)} is an algorithm to provide a single 
    number score to measure the performance of an SDO. 
    It works by statistically analyzing the past performance of the 
    organization and using that information to score an organization 
    on current performance.  CRI is a collection of the following 5 
    elements, which are actually KRIs,
    for an SDO.
    \textbf{
    \begin{enumerate}
        \item Quality
        \item Availability
        \item Satisfaction
        \item Schedule
        \item Requirements
    \end{enumerate}
    }
    A separate CRI
    score is calculated for each element and 
    then aggregated together to form an overall 
    CRI score.  A new CRI score will be calculated based
    upon the selection of a given time period (weekly, monthly, quarterly).
    CRI is not meant to be comparative between
    organizations, but to measure the amount of increase or decrease
    a single organization exhibits across elements.  CRI is made to be easily 
    expandable to other elements if desired.

@@

                The development and testing effort can come from any of 
                the following choices for effort. It is possible
                that other measures will work for effort.
                \begin{description}
                    \item[Actual Time] This number is a representation of the 
                        total amount of time spent on a project.  This number
                        can be measured in any unit of time: hours, days, weeks,
                        etc.  Actual time can be applied to development or testing
                        effort.
                    \item[Estimated Time] This number is a representation of the 
                        initial estimated amount of time spent on a project.  This number
                        can be measured in any unit of time: hours, days, weeks,
                        etc.  Estimated time can be applied to development or testing
                        effort.  It is common for the estimated and actual times
                        to be diffent.
                    \item[SLOC]  This number is the count of 
                        the total number of lines of source code for a project. 
                        Obviously, 
                        this item only counts as a level of
                        effort for development unless coding in used to generate
                        automated test cases. \footnote{Automated testing is the
                        process of creating software to automatically run tests
                        against other software.  The adoption of automated testing
                        is varied and it is not a solution in all cases 
                        \cite{Ramler2006}. }
                    \item[Modified Lines Of Code] This number is a count of the number 
                        of modified lines of source code. Modified lines is 
                        defined as the number of deleted, added, and modified 
                        lines of source code.  This number is different from above
                        since it does not include all the lines of source code.  Similar
                        to above, this number makes more sense for development effort.
                    \item[Test Cases] A test case is a step or series of steps followed
                        to validate some expected outcome of software.  Organizations
                        will create a number of test cases to be validated for a
                        software system.  The number of such test cases could
                        be used as a level of testing effort.
                \end{description}

### 2015-03-08 paper_50_sdlc-ae.tex
added abbreviations page

    In order for an SDO to properly track the elements of CRI, a data storage 
    system should be available to store the appropriate data.  A consistent 
    storage system should help to avoid the problem of inaccurate
    data caused by numerous manipulations of the existing data 
    \cite{Olson2003}. Plus, if the system
    is implemented correctly by allowing limited changes to existing data, 
    it will be able to alleviate
    some of the dishonesty that is currently present in software projects 
    \cite{Rost2011}. This storage system will be named the SDLC Analytic 
    Engine (SDLC-AE).

@@

    \subsection{DATABASE STRUCTURE}
        All the data necessary to compute CRI needs to be stored in a database.  This section will 
        lay out the structure of tables and relationships necessary to the store all the data
        within a relational database.  The SQL\footnote{Structured Query Language (SQL) is 
        a programming language designed for managing data in relational database 
        management system.} 
        is written for an Oracle
        database, but the scripts can be modified to work with another database
        such as: PostgreSQL, SQL Server, or MySQL.

>>>>

    In order for an SDO to properly track the elements of CRI, a data storage 
    system should be available to store the appropriate data.  A consistent 
    storage system should help to avoid the problem of inaccurate
    data caused by numerous manipulations of the existing data 
    \cite{Olson2003}. Plus, if the system
    is implemented correctly by allowing limited changes to existing data, 
    it will be able to alleviate
    some of the dishonesty that is currently present in software projects 
    \cite{Rost2011}. This storage system will be named the SDLC Analytic 
    Engine (SDLC-AE). \nomenclature{SDLC-AE}{SDLC Analytic Engine}

@@

    \subsection{DATABASE STRUCTURE}
        All the data necessary to compute CRI needs to be stored in a database.  This section will 
        lay out the structure of tables and relationships necessary to the store all the data
        within a relational database.  The SQL\footnote{Structured Query Language (SQL) is 
        a programming language designed for managing data in relational database 
        management system.\nomenclature{SQL}{Structured Query Language}} 
        is written for an Oracle
        database, but the scripts can be modified to work with another database
        such as: PostgreSQL, SQL Server, or MySQL.
